Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2021-10-11 methods_paper/fabl-paper.tex
new edits

%{
%	\color{blue} Thanks Beka for writing that first abstract. I think this one captures the idea better, even if its not exactly in "outline" form.
%	\vspace{1em}
%}

@@

{
	\color{blue}

@@

	Record linkage is the task of identifying duplicate records across multiple data sources, often in the absence of a unique identifier \citep{christen_2012}. This is an increasingly important task in ``data cleaning," and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others. In this paper, we consider bipartite record linkage, which is merging two databases that contain duplication across the two databases, but do not contain any duplications within each database.

@@

	In this paper, we propose the \emph{fast beta linkage} method (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space, creating a ``many-to-one" model for record linkage. This allows us to employ hashing techniques that hasten calculations and reduce computational costs. Furthermore, we can compute the pairwise record comparisons over large datasets via parallel computing and reduce memory costs through our proposed method of \emph{storage efficient indexing}. Throughout, we argue that even in cases where a bipartite matching is desired, \textcolor{red}{our proposed approach and relaxation} (with a simple post-processing procedure) provides equivalently accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed. We illustrate our results via simulation studies and a real case study of homicides from the El Salvadoran Civil War.

\textcolor{red}{BK: I would layout all the sections, section by section once they have been set for the reader and I can help you do this.}	
	% I think this is clear and has already been said. 
%	To recognize the lineage from the original \texttt{BRL} method, we name our method \emph{fast beta linkage} (\texttt{fabl}, pronounced ``fable'')

%	In what follows, we first review more thoroughly the Fellegi Sunter and Beta Record Linkage methods. Then, we provide the model specification under \texttt{fabl} and derive the Gibbs sampler for posterior inference. Next, we introduce the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of the linkage tasks we can undertake. Lastly, we use simulation studies and a case study of homicides from the El Salvadoran Civil War to show that our method has markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.

@@

Denote the set of matches by $\bm{M} = \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 1\}.$, and the set of non-matches by 
$\bm{U} =  \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying the sets of  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as "links", and record pairs that are estimated as non-matches as "non-links".

@@

Intuitively, records that are co-referent or rather refer to the same entity should be similar; records that are non co-referent and refer to different entities should not be similar. \cite{fellegi_theory_1969} proposed encoding this is using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote the number of criteria for comparing records by $F$, such that 
$\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots \gamma_{ij}^F).$ In most cases, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two datasets.

The simplest way to compare two records is check for agreement or disagreement, and this is commonly used for categorical variables. For more complex measurements, we can take into account partial agreement to more richly characterize the comparison; for numerical data, we can use absoluate difference, and for text data, we can use string distance metrics such as Levenstein or Jaro-Winkler distance. This allows us to divide our records into a set of similarity values into different levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}. Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f},$. Following convention, $I_{f0}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement). Thus, we can construct comparison vectors in the following way: 
$$\gamma_{ij}^f = \ell \; \text{if} \; \mathcal{S}_f(i,j) \in I_{f\ell}.$$

@@

One criticism of using mixture models is that the decision rule will lead to "many-to-many assignments," breaking one-to-one matching assumptions. One way of resolving this issue to incorporate the constraint directly into the model as follows:
%our model can be written as a Bayesian model for comparison data. In general, it has the same form of \cite{sadinle_bayesian_2017} and can be written as follows:
\begin{align}
	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m})  \\
	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag \\
	& \bm{Z} \sim \mathcal{B}, \notag
\end{align}
where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m}), \mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches. For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the \emph{beta distribution for bipartite matching}, given by
$$P(\mathbf{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\mathbf{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\mathbf{Z}) + \alpha_{\pi}, n_2 - \mathbf{Z}) - \beta_{\pi}])}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
$$
where $\text{B}(\cdot, \cdot)$ represents the Beta function, and hyper parameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$. This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate
records when sampling \(Z_j\). This creates a dependency that makes the
sampler \emph{inherently serial}, which makes the sampler slow when working on large linkage tasks.

@@

In practice, it is not feasible to make all-to-all record comparisons as the computationally complexity is of the order $O(n_1 \times n_2).$ The most common solution is utilize blocking, which places similar records into partitions, or "blocks" to reduces this computational burden. In deterministic blocking, the modeller chooses a field thought to be highly reliable, and only compares records that agree on that field. The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains. However, blocking on an unreliable field can lead to many missed matches, making this form of blocking undesirable in many situations.

@@

%------------------------------------------------------
%
%\textcolor{black}{I think its best just not to mention probabalistic blocking. My method doesn't use it, and introducing it is really confusing (Sadinle 2017, Enamorado 2019, and Steorts 2014/2015 don't mention it).}
%
%We note here that the word "hashing" is also used in probabalistic blocking, which uses techniques from machine learning to cluster records together before attempting to link records. The use of such probabalistic blocking to provide even greater computational gains will be explored in future work. 
%
%-------------------------------------------------------

@@

Importantly, this independence means that we have relaxed one-to-one requirement from \texttt{BRL}; our sampler does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\). This behavior may be desirable in its own right; \cite{Newcombe1962b} was interested in matching birth records to marriage certificates, a situation in which such "many-to-one" matchings are reasonable. In situations when a bipartite estimate is desired, we can use a simple post processing procedure shown in \ref{bayes-estimate} to resolve illegal matchings.

@@

\textcolor{black}{NOTE: I really want to present the entire model in one place like this. I saw this for the first time in Steorts 2014, and I really liked it. Part of the reason it took me so long to understand Sadinle 2017 was because the model was presented over several sections}

@@

Since our Gibbs procedure does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$. It is often desirable to see such matches, because they provide evidence of model misspecification; specifically, large numbers of "many to one" matches indicate that the assumption of no duplicates within files may not be reasonable. We will explore this issue more deeply through the El Salvador homicide case study.

@@

\section{Efficient and Scalable Implementation}
\label{sec:efficiency}
In this section, we propose efficient and scalable proposals using hashing based methods that are similar in spirit to \citep{enamorado2019}, a fast and scalable implementation of the Fellegi-Sunter model. 
 Specifically, one way we can improve our computational efficiency is by recognizing that record pairs contribute to posterior calculations only
through the agreement pattern of the \(\gamma_{ij}\) vector. To make this more precise, let 
\(\mathcal{H}\) be the set of unique agreement patterns in the data, let
\(P = |\mathcal{H}|\) denote the total number of unique agreement
patterns.  Observe that \(P\) is bounded above by \(\prod_{f=1}^F L_f\), and
that this bound does not depend on \(n_1\) or \(n_2\). Prior to
processing the data, we identify all \(P\) patterns in \(\mathcal{H}\)
and enumerate them as follows: \(h_1, \ldots, h_P\), which are called \emph{hashed values.}
Next, we map record pairs to their corresponding hashed value. 
More specifically, when the record pair \((i,j)\) 
exhibits the \(p^{th}\) agreement pattern, we say \((i,j) \in h_p\).
Whenever possible, we conduct calculations over these \(P\) agreement
patterns, instead of the typical  \(n_1 \times n_2\) record pairs.

%Some of these techniques are similar to those used by \citep{enamorado2019} to produce \texttt{fastlink}, a fast a scalable implementation of the Fellegi-Sunter model, and others have been created here for our particular context.

@@

In this section, we describe the hashing function we propose.

First, we hash record pairs of the same agreement pattern to unique integer values. \citep{enamorado2019} accomplished this efficiently through the hashing function
$$\tilde{\gamma}_{ij} = \sum_{f = 1}^F I(\gamma_{ij}^f > 0)2^{\gamma_{ij}^f + I(k>1)\sum_{e=1}^{k-1}(L_e -1)}$$
This function maps each agreement pattern to a unique integer, allowing us to store a scalar quantity instead of an entire vector for each record pair. For computational ease, we then map these integers to sequential integers from $1, \ldots, P.$

The classic Fellegi Sunter method represents the \(\gamma_{ij}\)
comparison as a vector of length \(F\), with each component
\(\gamma_{ij}^f\) taking on values in \(\{1, \ldots, L_f \}\). 
To ease our computational burden, we instead use a \emph{one hot encoding} of the comparison vector. For example, if
\(L_1 = L_2 = 2\) and \(L_3 = 3\), then \(\gamma_{ij} = (2, 1, 3)\)
under the classical framework becomes
\(\gamma_{ij} = (0, 1, 1, 0, 0, 0, 1)\) under our framework. This is a
bijective transformation that does not change the meaning of the data,
but this representation eases calculations and posterior updates. This is also the form the data takes in the \texttt{BRL} package in \texttt{R}.

In the classic Fellegi Sunter framework, $\bm{\Gamma}$ is a
\(n_1 n_2 \times F\) matrix, with each row providing the comparison
vector for a different \((i,j)\) pair. In contrast, we do not store
these comparison vectors themselves, but instead only the hashed value $h_p$
corresponding to the agreement pattern of the \((i, j)\) pair. We store
this information in a nested list $\tilde{\bm{\Gamma}}$ where the
\(p^{th}\) component of the \(j^{th}\) list contains a vector of records
in \(\bm{X}_1\) that share agreement pattern \(p\) with record \(j \in \bm{X}_2\). For
each \(p\), we also calculate
\(H_p = \sum_{i=1}^{n_1}\sum_{j=1}^{n_2} \mathbf{1}_{(i,j) \in h_p}\),
the total instances of agreement pattern \(p\) throughout the data, and
also for each \(j\), we calculate
\(H_{p_j} = \sum_{i=1}^{n_1} \mathbf{1}_{{(i,j) \in h_p}}\) the
instances of agreement pattern \(p\) among the comparison vectors
between record \(j \in \bm{X}_2\) and each of the \(n_1\) records in \(\bm{X}_1\).

The hashing procedure described above considerably reduces the memory
needed to store the comparison information. That is, instead of storing
\(n_1 \times n_1\) comparison vectors, which are relatively long under
either the Fellegi Sunter or our modified framework, we only store the
\(P\) unique vectors, and then \(n_1 \times n_2\) scalar quantities
relating record pairs to those vectors. However, even storing these
\(n_1 \times n_2\) scalar labels can become burdensome with large data.
Worse, the overwhelming majority of these labels relate to record pairs
that are clear non-matches.

To address this issue, we propose a new method called \emph{storage	efficient indexing} (SEI). In standard indexing, one decides a certain criteria that they expect all true matching pairs to satisfy. In addition, one decides 
a priori to label any record pairs that do not meet that criteria as
non-matches. For example, one might only consider pairs with a certain
similarity score on a field deemed to be important (like first name), or
only pairs with exact matching on a specified number of fields. While
generally chosen to be be quite loose, establishing these criteria
requires knowledge of the problem and invites room for human error. 
To improve upon this, we propose to reduce the comparison space and reduce the storage requirements while avoiding the drawbacks of standard indexing.

Observe that all records of the same
agreement pattern have the same probability when sampling \(Z_j\).
Therefore we know that records belonging to an \(h_p\) such that
\(H_{p_j}\) is large are very unlikely to be sampled consistently enough
to be deemed a match through the Bayes estimate, even without
considering the form of the agreement pattern itself.

In SEI, rather than store all of these unlikely record labels, we choose
to store only a small number \(R\) of them. Posterior calculations still
attribute the appropriate weight to all records through the summary
statistics \(H_p\), and \(H_{p_j}\). Rather than storing
\(n_1 \times n_2\) record labels, SEI allows us to store at most
\(n_2 \times P \times R\) labels, regardless of how large \(n_1\) might be.

Lastly, for large data, we can partition the two datasets \(\bm{X}_1\) and
\(\bm{X}_2\) into smaller blocks \(\{\bm{X}_{1m}\}\) and \(\{\bm{X}_{2m}\}\) for more
manageable computations. On a single machine, we can read-in data
sequentially, conduct hashing, compress information through SEI, and
delete the original data from memory before continuing with the next
chunk of data. With multiple cores or multiple machines, this can be
done in parallel. Thus, the combination of hashing, SEI, and partitioning
allows us to conduct linkage tasks over much larger datasets.

@@

\underline{Updating $\mathbf{m}$ and $\mathbf{u}$:} After receiving
matching statuses from \(\mathbf{Z}\), the Sadinle method calculates
\(\alpha_{fl}(\mathbf{Z})\) and \(\beta_{fl}(\mathbf{Z})\) for each
field and level. This constitutes \(2 \times \sum L_f\) many summations
over \(n_A \times n_B\) quantities, and becomes computationally
burdensome with large data. In contrast, we recognize that each unique
agreement pattern contributes to the posterior \(\alpha(\mathbf{Z})\)
and \(\beta(\mathbf{Z})\) vectors in the same way. In fact, if we denote
\(H_p^m = \sum_{j=1}^{n_B} \mathbf{1}_{(Z_j, j) \in h_p}\) to be the
number of matching record pairs with agreement pattern \(p\), then the
contribution of pairs of pattern \(p\) to the \(\alpha(\mathbf{Z})\)
vector is simply \(H_p^m \times h_p\). Thus our posterior update for the
\(\alpha\) vector is simply
\(\alpha(\mathbf{Z}) = \alpha_0 + \sum_{p=1}^P H_p^m \times h_p\). Then,
we can easily calculate \(H_p^u\), the number of nonmatching record
pairs of agreement pattern \(p\), by subtracting the number of matching
pairs from the total present in the data; that is
\(H_p^u = H_p - H_p^m\). From this, we can update our \(\beta\)
parameter through
\(\beta(\mathbf{Z}) = \beta_0 + \sum_{p=1}^P H_p^u \times h_p\). Note
that these constitute \(P\) many summations over \(n_B\) quantities, and
thus avoid the \(n_A \times n_B\) summation from the original method.

\underline{Updating $\mathbf{Z}$:} Although sampling \(Z_j\) from a the
full conditional provided earlier is conceptually straightforward, it
becomes computational burdensome when \(n_A\) is larger. The reader can
confirm that sampling a value from a large set of unequal probabilities
becomes difficult in most programming languages. To speed up
computation, we break this sampling step into two simpler steps. First,
we calculate the Fellegi Sunter weight \(w_{p}\) associated with each
unique pattern and sample the agreement pattern between \(j\) and its
potential match. Second, we sample the record label uniformly among
records associated with that agreement pattern for that particular
\(j\in B\). More concretely, define \(h(Z_j)\) to be the agreement
pattern between \(j\) and its potential match, and say
\(h(Z_j) = h_{P+1}\) when \(Z_j = n_A + 1\). Then,

>>>>

@@

@@

	Record linkage is the task of identifying duplicate records across multiple data sources, often in the absence of a unique identifier \citep{christen_2012}. This is an increasingly important task in ``data cleaning,'' and is used for inferential and predictive analyses in fields such as statistics, computer science, machine learning, political science, economics, precision medicine, official statistics, and others. In this paper, we consider bipartite record linkage, which is merging two databases that contain duplication across the two databases, but do not contain any duplications within each database.

@@

	In this paper, we propose the \emph{fast beta linkage} method (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. Specifically, we relax the one-to-one matching requirement of the Beta Record Linkage method of \citep{sadinle_bayesian_2017} and propose independent priors over the matching space, creating a ``many-to-one'' model for record linkage. This allows us to employ hashing techniques that hasten calculations and reduce computational costs. Furthermore, we can compute the pairwise record comparisons over large datasets via parallel computing and reduce memory costs through our proposed method of \emph{storage efficient indexing}. Throughout, we argue that even in cases where a bipartite matching is desired, \textcolor{red}{our proposed approach and relaxation} (with a simple post-processing procedure) provides equivalently accurate estimation of the linkage structure and other parameters, more information through which to asses model misspecification, and greatly enhanced speed.

	In what follows, Section \ref{sec:review-of_prior-work} reviews more thoroughly the Fellegi Sunter and Beta Record Linkage methods. Then, Section \ref{sec:fast-beta-linkage} provides the model specification under \texttt{fabl} and derives the Gibbs sampler for posterior inference. Next, Section \ref{sec:efficiency} introduces the hashing technique and storage efficient indexing used to increase the speed of calculations and the scale of the linkage tasks we can undertake. Lastly, in Sections \ref{sec:simulations} and \ref{sec:case-studies} we use simulation studies and a case study of homicides from the El Salvadoran Civil War to show that our method has markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}. We conclude in Section \ref{discussion} with a discussion of directions for future work.

	% I think this is clear and has already been said. 
%	To recognize the lineage from the original \texttt{BRL} method, we name our method \emph{fast beta linkage} (\texttt{fabl}, pronounced ``fable'')

@@

Denote the set of matches by $\bm{M} = \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 1\}.$, and the set of non-matches by 
$\bm{U} =  \{(i,j): i \in \bm{X}_1, j \in \bm{X}_2, \Delta_{ij} = 0\}.$ The record linkage task can be viewed as identifying the sets of  $\bm{M}$ and  $\bm{U}.$ We refer to record pairs that are estimated as matches as ``links'', and record pairs that are estimated as non-matches as ``non-links''.

@@

Intuitively, records that are co-referent or rather refer to the same entity should be similar; records that are non co-referent and refer to different entities should not be similar. \cite{fellegi_theory_1969} proposed encoding this is using a comparison vector $\bm{\gamma}_{ij}$ computed for each record pair $(i,j)$ in $\bm{X}_1 \times \bm{X}_2.$ Denote the number of criteria for comparing records by $F$, such that 
$\bm{\gamma}_{ij} = (\gamma_{ij}^1, \gamma_{ij}^2, \ldots, \gamma_{ij}^f, \ldots, \gamma_{ij}^F).$ In most cases, $\bm{\gamma}_{ij}$ consists of one comparison for each feature shared between the two datasets.

The simplest way to compare two records is check for agreement or disagreement, and this is commonly used for categorical variables. For more complex measurements, we can take into account partial agreement to more richly characterize the comparison; for numerical data, we can use absoluate difference, and for text data, we can use string distance metrics such as Levenstein or Jaro-Winkler distance. This allows us to divide our records into a set of similarity values into different levels of disagreement \citep{bilenko2006riddle, elmagarmid_duplicate_2007}. Let $\mathcal{S}_f(i,j)$ denote a general similarity measure for feature $f$ of records $i$ and $j,$ where the range of $\mathcal{S}_f$ can be divided into $L_f +1$ intervals denoted by $I_{f0}, I_{f1}, \ldots, I_{fL_f}$. Following convention, $I_{f0}$ represents the highest level of agreement (inclusive of complete agreement) and $I_{fL_f}$ represents the highest level of disagreement (including complete disagreement). Thus, we can construct comparison vectors in the following way: 
$$\gamma_{ij}^f = \ell \; \text{if} \; \mathcal{S}_f(i,j) \in I_{f\ell}.$$

@@

One criticism of using mixture models is that the decision rule will lead to ``many-to-many assignments,'' breaking one-to-one matching assumptions. One way of resolving this issue to incorporate the constraint directly into the model as follows:
%our model can be written as a Bayesian model for comparison data. In general, it has the same form of \cite{sadinle_bayesian_2017} and can be written as follows:
\begin{align}
	&\bm{\Gamma}_{ij} \mid Z_j =i \stackrel{iid}{\sim} \mathcal{M}(\bm{m})  \\
	&\bm{\Gamma}_{ij} \mid Z_j \neq i \stackrel{iid}{\sim} \mathcal{U}(\bm{u}) \notag \\
	& \bm{Z} \sim \mathcal{B}, \notag
\end{align}
where $\mathcal{B}$ represents a prior on the space of bipartite matchings, and $\mathcal{M}(\bm{m}), \mathcal{U}(\bm{u})$ are models for the comparison vectors of matches and non-matches. For \texttt{BRL}, \cite{sadinle_bayesian_2017} proposed the \emph{beta distribution for bipartite matching}, given by
$$P(\mathbf{Z}|\alpha_{\pi}, \beta_{\pi}) = \frac{(n_1 - n_{12}(\mathbf{Z}))!}{n_1 !}\frac{\text{B}(n_{12}(\mathbf{Z}) + \alpha_{\pi}, n_2 - \mathbf{Z}) - \beta_{\pi}])}{\text{B}(\alpha_{\pi}, \beta_{\pi})}
$$
where $\text{B}(\cdot, \cdot)$ represents the Beta function, and hyper parameters $\alpha_{\pi}$ and $\beta_{\pi}$ encode prior beliefs about the proportion of records in $\bm{X_2}$ that have matches in $\bm{X_1}$. This prior induces a Gibbs sampler that strictly enforces one-to-one matching, removing previously matched records from the set of candidate
records when sampling \(Z_j\). This creates a dependency that makes the
sampler \emph{inherently serial}, which makes the sampler slow when working on large linkage tasks.

@@

In practice, it is not feasible to make all-to-all record comparisons as the computationally complexity is of the order $O(n_1 \times n_2).$ The most common solution is utilize blocking, which places similar records into partitions, or ``blocks'' to reduces this computational burden. In deterministic blocking, the modeller chooses a field thought to be highly reliable, and only compares records that agree on that field. The record linkage method is then applied independently across all blocks, which can be done in parallel for additional speed gains. However, blocking on an unreliable field can lead to many missed matches, making this form of blocking undesirable in many situations.

@@

@@

Importantly, this independence means that we have relaxed one-to-one requirement from \texttt{BRL}; our sampler does ensure that each record in \(\bm{X}_2\) can be matched to at most one record in \(\bm{X}_1\), but allows for the possibility that multiple records in \(\bm{X}_2\) match to the same record in \(\bm{X}_1\). This behavior may be desirable in its own right; \cite{Newcombe1962b} was interested in matching birth records to marriage certificates, a situation in which such ``many-to-one'' matchings are reasonable. In situations when a bipartite estimate is desired, we can use a simple post processing procedure shown in \ref{bayes-estimate} to resolve illegal matchings.

@@

@@

Since our Gibbs procedure does not strictly enforce one-to-one matching, it is possible for this Bayes estimate to link multiple records in $\bm{X_2}$ to one record in $\bm{X_1}$. It is often desirable to see such matches, because they provide evidence of model misspecification; specifically, large numbers of ``many to one'' matches indicate that the assumption of no duplicates within files may not be reasonable. We will explore this issue more deeply through the El Salvador homicide case study.

@@

\section{Efficient and Scalable Implementation}
\label{sec:efficiency}

To increase the speed of our Gibbs sampler and expand the scale of the linkage tasks we can undertake, we employ hashing methods similar to those used by \cite{enamorado2019using} in the creation of \texttt{fastlink}, a fast and scalable implementation of the Fellegi Sunter method. The key insight is to recognize that record pairs contribute to posterior calculations only through the agreement pattern of the $\gamma_{ij}$ vector. To make this more precise, let 
\(\mathcal{H}\) be the set of unique agreement patterns in the data, let
\(P = |\mathcal{H}|\) denote the total number of unique agreement
patterns.  Observe that \(P\) is bounded above by \(\prod_{f=1}^F L_f\), and that this bound does not depend on \(n_1\) or \(n_2\).  Prior to
processing the data, we identify all \(P\) patterns in \(\mathcal{H}\)
and enumerate them \(1, \ldots, P\), which are called \emph{hashed values}. Additionally, we store ``one-hot encodings'' of these patterns rather than the original Fellegi Sunter comparison vectors, to aid in vectorized computations. Whenever possible, we conduct calculations over these \(P\) agreement
patterns, instead of the typical \(n_1 \times n_2\) record pairs.

%In this section, we propose efficient and scalable proposals using hashing based methods that are similar in spirit to \citep{enamorado2019}, a fast and scalable implementation of the Fellegi-Sunter model. 
% Specifically, one way we can improve our computational efficiency is by recognizing that record pairs contribute to posterior calculations only
%through the agreement pattern of the \(\gamma_{ij}\) vector. To make this more precise, let 
%\(\mathcal{H}\) be the set of unique agreement patterns in the data, let
%\(P = |\mathcal{H}|\) denote the total number of unique agreement
%patterns.  Observe that \(P\) is bounded above by \(\prod_{f=1}^F L_f\), and
%that this bound does not depend on \(n_1\) or \(n_2\). Prior to
%processing the data, we identify all \(P\) patterns in \(\mathcal{H}\)
%and enumerate them as follows: \(h_1, \ldots, h_P\), which are called \emph{hashed values.}
%Next, we map record pairs to their corresponding hashed value. 
%More specifically, when the record pair \((i,j)\) 
%exhibits the \(p^{th}\) agreement pattern, we say \((i,j) \in h_p\).
%Whenever possible, we conduct calculations over these \(P\) agreement
%patterns, instead of the typical  \(n_1 \times n_2\) record pairs.

@@

First, we hash record pairs of the same agreement pattern to unique integer values. \citep{enamorado2019using} accomplished this efficiently through the hashing function

$$h^{*}(i, j) = \sum_{f = 1}^F I(\gamma_{ij}^f > 0)2^{\gamma_{ij}^f + I(k>1)\sum_{e=1}^{k-1}(L_e -1)}$$

This function maps each agreement pattern to a unique integer, allowing us to store a scalar quantity instead of an entire vector for each record pair. For computational ease, we then map these integers to sequential integers from $1, \ldots, P$ corresponding to the enumerated patterns $\mathcal{H}$. When the $(i, j)$ pair exhbits the $p^{th}$ pattern, we say $h(i,j) = p$.

With all record pairs converted to hashed values, we can create a more compact nested list 
$$\mathcal{R} = \left\{\{r_{j_p} \}_{p=1}^{P} \right\}_{j = 1}^{n_2}$$
where $r_{j_p} = \{i \in \bm{X_1} | h(i,j) = p\}$. Note that the pair $(\mathcal{H}, \mathcal{R})$ fully characterizes the comparison matrix $\Gamma$ with no loss of information. This representation is useful because it allows us to easily compute sufficient statistics
$$H_{j_p} = \sum_{i} I(h(i,j) = p) = ||r_{j_p}||$$
and
$$H_p =  \sum_{i, j} I(h(i,j) = \sum_{j} H_{j_p}$$
As will be shown in Section \ref{efficient-posterior}, all posterior calculations are conducted through these sufficient statistics.

\hypertarget{large-data}{%
	\subsection{Working with Large Data}\label{large-data}}

The hashing procedure described above considerably reduces the memory
needed to store the comparison information. That is, instead of storing
\(n_1 \times n_2\) comparison vectors, we only store the
\(P\) unique vectors, and then \(n_1 \times n_2\) scalar quantities
relating record pairs to those vectors. However, even storing these
\(n_1 \times n_2\) scalar labels can become burdensome with large data.
Worse, the overwhelming majority of these labels relate to record pairs
that are clear non-matches.

%The classic Fellegi Sunter method represents the \(\gamma_{ij}\)
%comparison as a vector of length \(F\), with each component
%\(\gamma_{ij}^f\) taking on values in \(\{1, \ldots, L_f \}\). 
%To ease our computational burden, we instead use a \emph{one hot encoding} of the comparison vector. For example, if
%\(L_1 = L_2 = 2\) and \(L_3 = 3\), then \(\gamma_{ij} = (2, 1, 3)\)
%under the classical framework becomes
%\(\gamma_{ij} = (0, 1, 1, 0, 0, 0, 1)\) under our framework. This is a
%bijective transformation that does not change the meaning of the data,
%but this representation eases calculations and posterior updates. This is also the form the data takes in the \texttt{BRL} package in \texttt{R}.

%In the classic Fellegi Sunter framework, $\bm{\Gamma}$ is a
%\(n_1 n_2 \times F\) matrix, with each row providing the comparison
%vector for a different \((i,j)\) pair. In contrast, we do not store
%these comparison vectors themselves, but instead only the hashed value $h_p$
%corresponding to the agreement pattern of the \((i, j)\) pair. We store
%this information in a nested list $\tilde{\bm{\Gamma}}$ where the
%\(p^{th}\) component of the \(j^{th}\) list contains a vector of records
%in \(\bm{X}_1\) that share agreement pattern \(p\) with record \(j \in \bm{X}_2\). For
%each \(p\), we also calculate
%\(H_p = \sum_{i=1}^{n_1}\sum_{j=1}^{n_2} \mathbf{1}_{(i,j) \in h_p}\),
%the total instances of agreement pattern \(p\) throughout the data, and
%also for each \(j\), we calculate
%\(H_{p_j} = \sum_{i=1}^{n_1} \mathbf{1}_{{(i,j) \in h_p}}\) the
%instances of agreement pattern \(p\) among the comparison vectors
%between record \(j \in \bm{X}_2\) and each of the \(n_1\) records in \(\bm{X}_1\).

To address this issue, we propose a new method called \emph{storage	efficient indexing} (SEI). In standard indexing, one decides a priori certain criteria that they expect all true matching pairs to satisfy, 
and labels any record pairs that do not meet that criteria as
non-matches. For example, one might only consider pairs with a certain
similarity score on a field deemed to be important (like first name), or
only pairs with exact matching on a specified number of fields. While
generally chosen to be quite loose, establishing these criteria
requires knowledge of the problem and invites room for human error.

With SEI however, we can reduce comparison space, and its associated storage costs, while avoiding these drawbacks. Observe that all records of the same agreement pattern have the same probability when sampling \(Z_j\). Therefore we know that records $\i in r_{j_p}$  such that
\(H_{j_p}\) is large are very unlikely to be sampled consistently enough
to be deemed a match through the Bayes estimate. We know this regardless of the form of the agreement pattern itself, or its associated probabilities. Therefore, rather than store all of these unlikely record labels, we choose to store only a small number \(S\) of them in a new nested listed $\mathcal{R}^{\text{SEI}}$. Posterior calculations still attribute the appropriate weight to all records through the summary statistics \(H_p\), and \(H_{j_p}\). Rather than storing
\(n_1 \times n_2\) record labels, SEI allows us to store at most
\(n_2 \times P \times S\) labels, regardless of how large \(n_1\) might be.

Lastly, for large data, we can partition the two datasets \(\bm{X}_1\) and
\(\bm{X}_2\) into smaller blocks \(\{\bm{X}_{1m}\}\) and \(\{\bm{X}_{2n}\}\) for more
manageable computations. On a single machine, we can read-in data
sequentially, conduct hashing, compress information through SEI, and
delete the original data from memory before continuing with the next
chunk of data. With multiple cores or multiple machines, this can be
done in parallel. Summary statistics from each pairwise chunk comparison can then be easily synthesized to recover sufficient statistics for the larger linkage task.  Thus, the combination of hashing, SEI, and partitioning
allows us to conduct linkage tasks over much larger datasets.

@@

\underline{Sample $\Phi^{(s+1)}|\mathcal{H}, \mathcal{R}, \mathbf{Z}^{(s)}$:} Calculated at the level of the record pairs, updating  \(\alpha_{fl}(\mathbf{Z})\) and \(\beta_{fl}(\mathbf{Z})\) for each field and level in the linkage task constitutes \(2 \times \sum L_f\) many summations over \(n_A \times n_B\) quantities. These are simple calculations, but become computationally burdensome when working on large linkage tasks.

Instead, we use one-hot encodings of the agreement patterns $\mathcal{H}$ for more efficient calculations. Denote
$H_p^m = \sum_{j=1}^{n_B} \mathbf{1}_{(Z_j, j) \in h_p}$ to be the
number of matching record pairs with agreement pattern $p$. We can also easily calculate the number of nonmatching record pairs with agreement pattern $p$ as \(H_p^u = H_p - H_p^m\). Then, if $\mathbf{\alpha_0}$ and $\mathbf{\beta_0}$ are vectors of prior parameters for the $\mathbf{m}$ and $\mathbf{u}$ distributions respectively, the posterior update becomes simply

$$\alpha(\mathbf{Z}) = \mathbf{\alpha_0} + \sum_{p=1}^P H_p^m \times h_p$$
$$\beta(\mathbf{Z}) = \mathbf{\beta_0} + \sum_{p=1}^P H_p^u \times h_p$$

Note that these constitute \(P\) many summations over \(n_B\) quantities, and thus avoid the \(n_A \times n_B\) summation from the original method.
